---  
title: "RHOGE"  
author: "Nick Mancuso"  
date: "`r format(Sys.Date())`"
output: github_document
---  
RHOGE is an R package that computes estimates the genome-wide genetic correlation between two complex traits (diseases) as
a function of predicted gene expression effect on trait (\\rho_{ge}). Given output from two transcriptome-wide association studies,
RHOGE estimates the mediating effect of predicted gene expression and estimates the correlation of effect sizes. This approach
is extended to a bi-directional regression that provides putative causal directions between traits with non-zero \\rho_{ge}.

This approach is described in:

> [Integrating Gene Expression with Summary Association Statistics to Identify Genes Associated with 30 Complex Traits](https://doi.org/10.1016/j.ajhg.2017.01.031)\
> Nicholas Mancuso, Huwenbo Shi, Pagé Goddard, Gleb Kichaev, Alexander Gusev, Bogdan Pasaniuc.\
> American Journal of Human Genetics. 2017.

# Installation
Bioconductor + devtools is the most straightforward way to install RHOGE. To do this open an R terminal and enter
```{r eval=FALSE}
source("http://bioconductor.org/biocLite.R")
biocLite("devtools")    # only if devtools not yet installed
biocLite("bogdanlab/RHOGE")
```

# Example
The following example computes \\rho_{ge} between BMI and triglycerides, as well as putative causal directions.
```{r warning=FALSE}
library(RHOGE)

# example BMI TWAS results from FUSION
data(bmi)
head(bmi)

# example triglyceride TWAS results from FUSION
data(tg)
head(tg)

# Estimate rho_ge genome-wide for BMI and Triglyerides and approximate sample sizes
ge_cor_res <- rhoge.gw(bmi, tg, 14000, 91000)
head(ge_cor_res)

bidir_res <- rhoge.bd(bmi, tg, 14000, 91000, p1 = 0.05 / nrow(bmi), p2 = 0.05 / nrow(tg))
head(bidir_res)
```

Currently, only [FUSION](https://github.com/gusevlab/fusion_twas) style output is supported.
